  Reproducible Research Project 1
===================================


  Code for reading in the dataset and/or processing the data
```{r echo = TRUE}
library(dplyr)
library(ggplot2)
activity <- read.csv("activity.csv")
summary(activity)
```
Processing the data
```{r echo=TRUE}
activity$day <- weekdays(as.Date(activity$date))
```

Calculated the total number of steps taken per day
```{r echo = TRUE}
sumTable <- aggregate(activity$steps ~ activity$date, FUN=sum)
colnames(sumTable)<- c("Date", "Steps")
```

 Histogram of the total number of steps taken each day
```{r echo = TRUE}
hist(sumTable$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day",col="grey")
```

 Calculate the mean and median for the number of steps taken each day
```{r echo = TRUE}
## Mean for the number of steps each day
round(mean(sumTable$Steps))
## Median for the number of steps each day
round(median(sumTable$Steps))
```

The average number of steps taken
```{r echo = TRUE}
avg_steps <- aggregate(activity$steps ~ activity$interval, FUN=mean)
colnames(avg_steps)<- c("interval","Average_Steps")
```

 Time series plot
```{r echo = TRUE}
plot(avg_steps$interval,avg_steps$Average_Steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

The maximum number of steps taken in 5-minute interval
```{r echo =TRUE}
avg_steps$Average_Steps <- as.integer(avg_steps$Average_Steps)
max <- max(avg_steps$Average_Steps)
max_step <-avg_steps[avg_steps$Average_Steps==max,1]
```
The interval `r max_step` having `r max` number of step 

Imputing missing values
```{r echo = TRUE}
nadata<- activity[is.na(activity$steps),]
## Number of rows having missing values
nrow(nadata)
##Merge NA data with average steps data
newdata<-merge(activity, avg_steps, by= "interval")
## Inset zero to first day
newdata[as.character(newdata$date) == "2012-10-01", 2] <- 0
## arrange date wise
newdata <- arrange(newdata, date)
##Replace NA with average steps
newdata$steps[is.na(newdata$steps)] <- newdata$Average_Steps 
```

the total number of steps taken each day after missing values are imputed
```{r echo=TRUE}
newdata2 <- aggregate(newdata$steps ~ newdata$date, FUN=sum)
colnames(newdata2)<- c("date", "Steps")
```

Histogram of the total number of steps taken each day after missing values are imputed
```{r echo = TRUE}
hist(newdata2$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day", col="black")
hist(sumTable$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day",col = "grey",add = T)
legend("topright", c("Imputed", "Non-imputed"), col=c("black", "grey"), lwd=10)
```

```{r echo=TRUE}
newdata$DayCategory <- ifelse(newdata$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
```

Plot the data
```{r echo=TRUE}
plot<- ggplot(newdata, aes(x =interval , y=steps, color=DayCategory)) +
       geom_line() +
       labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") +
       facet_wrap(~DayCategory, ncol = 1, nrow=2)
print(plot)
```